DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

In \cite{kondor2018clebsch}, the convolution is performed in the spectral (harmonic) domain with a simple Hadamard product.
Instead of recomposing the signal and applying a traditional non-linearity such as the ReLU in the pixel domain, \cite{kondor2018clebsch} proposes to apply a Clebsch–Gordan product in the spectral domain.
(As a non-linear operation preserving rotation equivariance, it is considered as the non-linearity of the NN.)
Their convolution costs $\mathcal{O}(N_{pix})$, and their non-linearity costs $\mathcal{O}(N^{3/2}_{pix})$.
Stacking those layers result in a NN of overall $\mathcal{O}(N^{3/2}_{pix})$ time complexity.
We clarified the sentence.

\subsection{}
\begin{mdframed}[style=comment]
When mentioning that you are 10 to 20 times faster than the SHT it should be noted that libsharp has amazing distributed mpi functionality which helps, although of course your method is genuinely lower order computationally (and can be distributed simply on a GPU).
\end{mdframed}

True.
In the experiment, both were compared on a single core.
They might however scale differently given one can be distributed through MPI and the other on a GPU.
The manuscript was updated.

\subsection{}
\begin{mdframed}[style=comment]
In Appendix A. (my personal favourite part) could you measure the percent difference that different eigenvectors differ from the spherical harmonics due to incomplete discretisation of the sphere, perhaps plotting side by side plots like Figure 3, but with SH too (or their differences). By the way, Figure A.12 is placed in the wrong section.
\end{mdframed}

% (To plot the difference between the graph eigenvectors and the spherical harmonics on the sphere, one would have to align the subspaces.) Impossible to rotate arbitrarily on HEALPix.
We cannot compare individual eigenvectors and spherical harmonics because of arbitration rotations.
While any rotation of the basis of a subspace (the space spanned by the eigenvectors and harmonics corresponding to a given degree $\ell$) doesn't alter the spanned subspace, the difference between the basis vectors before and after the rotation can be 100\%.

Your suggestion however made us realize that figure A.12 was not ideal in comparing the eigenvectors to the spherical harmonics.
As such, we replaced it with another that we hope is clearer.
We also moved Figure A.12 in the correct subsection.

@@

\begin{mdframed}[style=comment]
The output’s size of the neural network depends on the task.

- It would be more correct to write "The size of the output of the neural network depends on the task."
\end{mdframed}
We agree with the reviewer and have updated the manuscript.

\begin{mdframed}[style=comment]
Its radius should be large enough to capture statistics of interest.

- I think you mean the radius of the kernel. Perhaps this could be rewritten.
\end{mdframed}
% \todo{Assign: @tomek}
% \todo{To check: @all}
% \nati{The manuscript seems clear to me. But let me know if you think otherwise.}
We updated the text slightly.

### 2019-03-04 scnn.tex
comment 2.11

\section{Graph Fourier modes and spherical harmonics}
\label{sec:comparison_spherical_harmonics}
The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}.
Let us further observe the spectral properties of our constructed spherical graph laplacian $\L$.
Its eigenvalues, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
We remind the reader that the amplitude of the Laplacian eigenvalue is proportional to the sum of the variations of its associated eigenvector. Furthermore, all spherical harmonics with the same order $\ell$ have the same variation. Hence, the fact that the Laplacian eigenvalues are grouped in blocks of size $2\ell + 1$ is a hint that the graph eigenvectors approximate the spherical harmonics.
In order to push the comparison one step further, we show the correspondence (scalar product) between the subspaces spanned by the graph Fourier modes and the spherical ha, however, out of the scope of this contribution.rmonics in \figref{subspace_harmonics_eigenvectors}.
The block-diagonal structure of the matrix, as opposed to diagonal, results the fact that the phase of the graph eigenvector is not aligned with the one of the spherical harmonics.

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking outside of the block diagonal in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
% We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
The theoretical study of the convergence is, however, out of the scope of this contribution.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, which corresponds to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell + 1$ orders. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors}
	\includegraphics[width=\linewidth]{figures/subspace_alignment_percentage.pdf}
	\caption{Top: Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics. Each pixel on the plot is the dot product between one spherical harmonic and one graph Fourier mode $\b u_i$. The plot clearly shows how the subspaces are aligned: modes and harmonics that correspond to different degrees $\ell$ are mostly orthogonal to each other, while modes and harmonics that correspond to the same degree $\ell$ are linearly dependent. There are $2 \ell + 1$ orders per degree $\ell$. Moreover, we see that the Fourier modes have the correct symmetries, indicated by the perfect orthogonality between even and odd functions. For example, the modes corresponding to $\ell=7$ are perfectly orthogonal to the harmonics corresponding to $\ell=0,2,4,6,8$.
	\nati{Bottom: Evolution of the allignement of the graph Fourier mode and the spherical harmonics for with respect of the speherical harmonic index $\ell$. It is the percentage of energy contained in the diagonal blocks of the correlation matrix plotted on top.}}
	\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

>>>>

\section{Graph Fourier modes and spherical harmonics}
\label{sec:comparison_spherical_harmonics}

The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}.
Let us further observe the spectral properties of our constructed spherical graph laplacian $\L$.
Its eigenvalues, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
We remind the reader that the amplitude of the Laplacian eigenvalue is proportional to the sum of the variations of its associated eigenvector. Furthermore, all spherical harmonics with the same order $\ell$ have the same variation. Hence, the fact that the Laplacian eigenvalues are grouped in blocks of size $2\ell + 1$ is a hint that the graph eigenvectors approximate the spherical harmonics.
In order to push the comparison one step further, we show the correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics in \figref{subspace_harmonics_eigenvectors}.
%The block-diagonal structure of the matrix, as opposed to diagonal, results the fact that the phase of the graph eigenvector is not aligned with the one of the spherical harmonics.

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, which corresponds to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell + 1$ orders. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors_v2}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics.
		% No real convergence with increasing Nside.
		Row $j$ is the sum of the (normalized) power spectral densities (PSDs), computed by the SHT, of the $2j+1$ graph Fourier modes $\b u_i$ identified with degree $j$.
		(\figref{graph_eigenvalues} shows how Fourier modes can be identified to spherical degrees.)
		The PSD is computed for degrees $\ell=0$ to $\ell = 3 N_{side} -1$.
		Pixel $k$ in row $j$ measures the energy in frequency band $k$ of the graph Fourier modes identified to be part of band $j$.
		The matrix should ideally be the identity.
		While not the case, the plot shows how the subspaces agree: the subspaces spanned by the Fourier modes correspond to those spanned by the spherical harmonics in the low frequencies, while the Fourier modes leak towards adjacent frequency bands in the high frequencies.
		While there is a systematic error, the Fourier modes agree at higher frequencies as $N_{side}$ increases.}
		\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking across frequency bands in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
% We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
The theoretical study of the convergence is, however, out of the scope of this contribution.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.
